---
title: "Business Intelligence Lab Submission Markdown"
author: "Team Marafiki"
date: "23/09/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+---------------------------------------------+
| **Student ID Numbers and Names of Group Members** | 1.  136446 - C - Mirav Bhojani              |
|                                                   |                                             |
|                                                   | 2.  136788 - C - Derrick Nyaga              |
|                                                   |                                             |
|                                                   | 3.  136709 - C - Jane Mugo                  |
|                                                   |                                             |
|                                                   | 4.  136895 - C - Wesley Wanyama             |
|                                                   |                                             |
|                                                   | 5.  135399 - C - Sheilla Kavinya            |
+---------------------------------------------------+---------------------------------------------+
| **GitHub Classroom Group Name**                   | *Team Marafiki*                             |
+---------------------------------------------------+---------------------------------------------+
| **Course Code**                                   | BBT4206                                     |
+---------------------------------------------------+---------------------------------------------+
| **Course Name**                                   | Business Intelligence II                    |
+---------------------------------------------------+---------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology |
+---------------------------------------------------+---------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023  |
+---------------------------------------------------+---------------------------------------------+

# Setup Chunk

We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## formatR - Required to format R code in the markdown ----
if (!is.element("formatR", installed.packages()[, 1])) {
  install.packages("formatR", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("formatR")


## readr - Load datasets from CSV files ----
if (!is.element("readr", installed.packages()[, 1])) {
  install.packages("readr", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("readr")
```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

# Performance Dataset

The 20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset is then loaded.

```{r Load Dataset}
student_performance_dataset <-
  readr::read_csv(
                  "../data/20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset.CSV", # nolint
                  col_types =
                  readr::cols(
                              class_group =
                              readr::col_factor(levels = c("A", "B", "C")),
                              gender = readr::col_factor(levels = c("1", "0")),
                              YOB = readr::col_date(format = "%Y"),
                              regret_choosing_bi =
                              readr::col_factor(levels = c("1", "0")),
                              drop_bi_now =
                              readr::col_factor(levels = c("1", "0")),
                              motivator =
                              readr::col_factor(levels = c("1", "0")),
                              read_content_before_lecture =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              anticipate_test_questions =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              answer_rhetorical_questions =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              find_terms_I_do_not_know =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              copy_new_terms_in_reading_notebook =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              take_quizzes_and_use_results =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              reorganise_course_outline =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              write_down_important_points =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              space_out_revision =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              studying_in_study_group =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              schedule_appointments =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              goal_oriented =
                              readr::col_factor(levels =
                                                c("1", "0")),
                              spaced_repetition =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              testing_and_active_recall =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              interleaving =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              categorizing =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              retrospective_timetable =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              cornell_notes =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              sq3r = readr::col_factor(levels =
                                                       c("1", "2", "3", "4")),
                              commute = readr::col_factor(levels =
                                                          c("1", "2",
                                                            "3", "4")),
                              study_time = readr::col_factor(levels =
                                                             c("1", "2",
                                                               "3", "4")),
                              repeats_since_Y1 = readr::col_integer(),
                              paid_tuition = readr::col_factor(levels =
                                                               c("0", "1")),
                              free_tuition = readr::col_factor(levels =
                                                               c("0", "1")),
                              extra_curricular = readr::col_factor(levels =
                                                                   c("0", "1")),
                              sports_extra_curricular =
                              readr::col_factor(levels = c("0", "1")),
                              exercise_per_week = readr::col_factor(levels =
                                                                    c("0", "1",
                                                                      "2",
                                                                      "3")),
                              meditate = readr::col_factor(levels =
                                                           c("0", "1",
                                                             "2", "3")),
                              pray = readr::col_factor(levels =
                                                       c("0", "1",
                                                         "2", "3")),
                              internet = readr::col_factor(levels =
                                                           c("0", "1")),
                              laptop = readr::col_factor(levels = c("0", "1")),
                              family_relationships =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              friendships = readr::col_factor(levels =
                                                              c("1", "2", "3",
                                                                "4", "5")),
                              romantic_relationships =
                              readr::col_factor(levels =
                                                c("0", "1", "2", "3", "4")),
                              spiritual_wellnes =
                              readr::col_factor(levels = c("1", "2", "3",
                                                           "4", "5")),
                              financial_wellness =
                              readr::col_factor(levels = c("1", "2", "3",
                                                           "4", "5")),
                              health = readr::col_factor(levels = c("1", "2",
                                                                    "3", "4",
                                                                    "5")),
                              day_out = readr::col_factor(levels = c("0", "1",
                                                                     "2", "3")),
                              night_out = readr::col_factor(levels = c("0",
                                                                       "1", "2",
                                                                       "3")),
                              alcohol_or_narcotics =
                              readr::col_factor(levels = c("0", "1", "2", "3")),
                              mentor = readr::col_factor(levels = c("0", "1")),
                              mentor_meetings = readr::col_factor(levels =
                                                                  c("0", "1",
                                                                    "2", "3")),
                              `Attendance Waiver Granted: 1 = Yes, 0 = No` =
                              readr::col_factor(levels = c("0", "1")),
                              GRADE = readr::col_factor(levels =
                                                        c("A", "B", "C", "D",
                                                          "E"))),
                  locale = readr::locale())
```

## Description of the Dataset

We then display the number of observations and number of variables. We have 101 observations and 100 variables to work with.

```{r Display dimension}
dim(student_performance_dataset)
```

Next, we display the quartiles for each numeric variable

```{r Display data summary}
summary(student_performance_dataset)
```

# Measures of Frequency for Categorical Variables

This code chunk is used to calculate the frequency and percentage distribution of the categorical variables in the dataset.

```{r Frequency}
student_performance_freq <- student_performance_dataset$class_group
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$gender
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$regret_choosing_bi
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$drop_bi_now
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$motivator
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$read_content_before_lecture
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$anticipate_test_questions
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$answer_rhetorical_questions
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$find_terms_I_do_not_know
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$copy_new_terms_in_reading_notebook
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$take_quizzes_and_use_results
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$reorganise_course_outline
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$write_down_important_points
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$space_out_revision
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$studying_in_study_group
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$schedule_appointments
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$goal_oriented
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$spaced_repetition
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$testing_and_active_recall
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$interleaving
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$categorizing
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$retrospective_timetable
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$cornell_notes
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$sq3r
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$commute
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$study_time
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$repeats_since_Y1
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$paid_tuition
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$free_tuition
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$extra_curricular
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$sports_extra_curricular
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$exercise_per_week
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$meditate
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$pray
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$internet
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$laptop
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$family_relationships
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$romantic_relationships
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$spiritual_wellnes
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$financial_wellness
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$health
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$day_out
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$night_out
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$alcohol_or_narcotics
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$mentor
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$mentor_meetings
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$`A - 1. I am enjoying the subject`
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)

student_performance_freq <- student_performance_dataset$`A - 2. Classes start and end on time`
cbind(frequency = table(student_performance_freq),
      percentage = prop.table(table(student_performance_freq)) * 100)
```

## Measures of Central Tendency(Calculating the Mode)

This code chunk displays code used to calculate the mode of variable in the dataset. A function was manually created to calculate the mode.

```{r Mode}
student_performance_class_group_mode <- names(table(student_performance_dataset$class_group))[
  which(table(student_performance_dataset$class_group) == max(table(student_performance_dataset$class_group)))
]
print(student_performance_class_group_mode)

student_performance_gender_mode <- names(table(student_performance_dataset$gender))[
  which(table(student_performance_dataset$gender) == max(table(student_performance_dataset$gender)))
]
print(student_performance_gender_mode)

student_performance_regret_choosing_bi_mode <- names(table(student_performance_dataset$regret_choosing_bi))[
  which(table(student_performance_dataset$regret_choosing_bi) == max(table(student_performance_dataset$regret_choosing_bi)))
]
print(student_performance_regret_choosing_bi_mode)

student_performance_drop_bi_now_mode <- names(table(student_performance_dataset$drop_bi_now))[
  which(table(student_performance_dataset$drop_bi_now) == max(table(student_performance_dataset$drop_bi_now)))
]
print(student_performance_drop_bi_now_mode)

student_performance_motivator_mode <- names(table(student_performance_dataset$motivator))[
  which(table(student_performance_dataset$motivator) == max(table(student_performance_dataset$motivator)))
]
print(student_performance_motivator_mode)

student_performance_read_content_before_lecture_mode <- names(table(student_performance_dataset$read_content_before_lecture))[
  which(table(student_performance_dataset$read_content_before_lecture) == max(table(student_performance_dataset$read_content_before_lecture)))
]
print(student_performance_read_content_before_lecture_mode)

student_performance_anticipate_test_questions_mode <- names(table(student_performance_dataset$anticipate_test_questions))[
  which(table(student_performance_dataset$anticipate_test_questions) == max(table(student_performance_dataset$anticipate_test_questions)))
]
print(student_performance_anticipate_test_questions_mode)

student_performance_find_terms_mode <- names(table(student_performance_dataset$find_terms_I_do_not_know))[
  which(table(student_performance_dataset$find_terms_I_do_not_know) == max(table(student_performance_dataset$find_terms_I_do_not_know)))
]
print(student_performance_find_terms_mode)

student_performance_copy_new_terms_mode <- names(table(student_performance_dataset$copy_new_terms_in_reading_notebook))[
  which(table(student_performance_dataset$copy_new_terms_in_reading_notebook) == max(table(student_performance_dataset$copy_new_terms_in_reading_notebook)))
]
print(student_performance_copy_new_terms_mode)

student_performance_take_quizzes_mode <- names(table(student_performance_dataset$take_quizzes_and_use_results))[
  which(table(student_performance_dataset$take_quizzes_and_use_results) == max(table(student_performance_dataset$take_quizzes_and_use_results)))
]
print(student_performance_take_quizzes_mode)

student_performance_reorganize_course_outline_mode <- names(table(student_performance_dataset$reorganise_course_outline))[
  which(table(student_performance_dataset$reorganise_course_outline) == max(table(student_performance_dataset$reorganise_course_outline)))
]
print(student_performance_reorganize_course_outline_mode)
```

## Standard Deviation of each variable

This code chunk was used to calculate the standard deviation of each variable. The variables are categorical making meaning we had to convert the values to numerical values. This was done after selecting the columns which we wanted to calculate the standard deviation.

```{r Standard Deviation}
selected_columns <- student_performance_dataset[, c(1:100)]
selected_columns <- sapply(selected_columns, as.numeric)
standard_deviations <- sapply(selected_columns, sd)
print(standard_deviations)
```

## Variance of each Variable

This chunk displays code on how the variance was calculated using the above selected columns.

```{r Variance}
variance <- sapply(selected_columns, var)
print(variance)

```

## Kurtosis of each variable

This chunk displays how the kurtosis of each variable was calculated. Initially the moments library was installed then loaded to facilitate this. In the selected columns, the columns that have at least 4 non-missing values were filtered out before calculating the kurtosis.

```{r Kurtosis}
library(moments)

if (!is.element("e1071", installed.packages()[, 1])) {
  install.packages("e1071", dependencies = TRUE)
}
require("e1071")
selected_columns_filtered <- selected_columns[sapply(selected_columns, function(col) sum(!is.na(col)) >= 4)]
kurtosis <- sapply(selected_columns_filtered, kurtosis, type = 2)
print(kurtosis)

```

## Skewness of each variable

```{r Skewness}
skewness <- sapply(selected_columns, skewness)
print(skewness) 
```

## Covariance between variables

```{r Covariance}
student_performance_cov <- cov(selected_columns)
View(student_performance_cov) 
```

## Correlation between variables

```{r Correlation}
student_performance_cor <- cor(selected_columns)
View(student_performance_cor) 
```

## 
